A Comparative Quality Metric for Fuzzing with Logic State Coverage

### 2024-02-26 paper/main.tex
minor

  To establish a more reliable quality metric, we first note that fuzzing 
  aims to reduce the number of unknown \emph{abnormal} behaviors by observing
  more \emph{interesting} (i.e., relating to unknown abnormal) behaviors. This
  suggests that the number of observed interesting behaviors should directly
  indicate how well the fuzzer achieved its aim (i.e., the quality).

  In this work, we propose \emph{logic state coverage} as a proxy metric to
  count observed interesting behaviors. A logic state distinguishes interesting
  behaviors in a finer granularity than the less interesting ones, making the
  amount of logic state coverage (i.e., observed logic states) directly
  proportional to the number of observed interesting behaviors.

  We implemented logic state coverage using a bloom filter and reevaluated
  \textsc{FuzzBench} and indecisive-to-subjective fuzzer comparisons in
  hypervisor and grammar fuzzing. 
  % TODO: (The result summary will come here)

@@

\pp{Interestingness in practice}
%
As the interestingness $P(\mathrm{abnormal})$ depends on the expectation of an
\emph{unobserved} behavior, there is no direct way to measure or formalize
interestingness. However, the practice of untargeted fuzzing in recent years
suggests that abnormal behaviors get increasingly unlikely to require a
large number of branch hit when it gets larger.
%
The prime example is the branch hit count bucketing in AFL \cite{afl}, where the
bucket size gets larger roughly by the power of 2 (i.e., 1, 2, 3, 4-7, 8-15,
...).
%; all hit counts are treated the same within a bucket.
%
%This is also intuitive because if a certain abnormal behavior can manifest at a
%low loop count, such a behavior is not \emph{unknown} anymore by when it reaches
%a higher loop count.

@@

\autoref{e:count} comes with three major benefits. First, it offers one of the
best accuracy among all alternatives 
\cite{harmouch2017cardinality}. 
%, while its error rate is \emph{inversely proportional} to the actual number of
%elements in the set (i.e., even \emph{less} error as logic state coverage
%grows) \cite{harmouch2017cardinality}. 
%
Second, it allows for a near-constant-time calculation as it only requires the
number of 1s apart from the formula calculation, as well as empirically shown in
\cite{harmouch2017cardinality} (\textbf{R2}). 
%
Finally, it presents a calculable error bound that the actual number of elements
likely falls within given a confidence probability \cite{papapetrou2010cardinality}.

@@

\autoref{f:design} illustrates the workflow of logic state coverage
measurement. Similar to edge coverage, it first records and buckets the branch
hit counts of every execution, yielding a \emph{logic state} as a result 
(\autoref{s:design:rec}). Then, it adds the recorded logic state to a 
\emph{bloom filter}, representing a set of observed logic states, by setting the
hashed indices (\autoref{s:design:up}).  Finally, the measurement periodically
reads off the bloom filter to count the number of observed logic states
(\autoref{s:design:count}).

>>>>

  To establish a more reliable quality metric, we first note that fuzzing 
  aims to reduce the number of unknown \emph{abnormal} behaviors by observing
  more \emph{interesting} (i.e., relating to unknown abnormal) behaviors. The
  more interesting behaviors a fuzzer has observed, the stronger guarantee it
  can provide about the absence of unknown abnormal behaviors. This suggests
  that the number of observed interesting behaviors must directly indicate the
  fuzzing quality.

  In this work, we propose \emph{logic state coverage} as a proxy metric to
  count observed interesting behaviors. A logic state distinguishes interesting
  behaviors in a finer granularity than the less interesting ones, making the
  amount of logic state coverage (i.e., observed logic states) directly
  proportional to the number of observed interesting behaviors.
  % 
  We implemented logic state coverage using a bloom filter and reevaluated
  \textsc{FuzzBench} and indecisive-to-subjective fuzzer comparisons in
  hypervisor and grammar fuzzing. 
  % TODO: (The result summary will come here)

@@

\pp{Interestingness in practice}
%
As the interestingness $P(\mathrm{abnormal})$ depends on the expectation of an
\emph{unobserved} behavior, there is no direct way to measure or formalize
interestingness. However, the practice of untargeted fuzzing in recent years
suggests that abnormal behaviors get increasingly unlikely to require a
larger number of branch hit.
%
The prime example is the branch hit count bucketing in AFL \cite{afl}, where the
bucket size gets larger roughly by the power of 2 (i.e., 1, 2, 3, 4-7, 8-15,
...).
%; all hit counts are treated the same within a bucket.
%
%This is also intuitive because if a certain abnormal behavior can manifest at a
%low loop count, such a behavior is not \emph{unknown} anymore by when it reaches
%a higher loop count.

@@

\autoref{e:count} comes with three major benefits. First, it offers one of the
best accuracy among all alternatives (c.f., only 4\% error even in an extreme
filter density of 90\%) \cite{harmouch2017cardinality}. 
%, while its error rate is \emph{inversely proportional} to the actual number of
%elements in the set (i.e., even \emph{less} error as logic state coverage
%grows) \cite{harmouch2017cardinality}. 
%
Second, it allows for a near-constant-time calculation as it only requires the
number of 1s apart from the formula calculation, as well as empirically shown in
\cite{harmouch2017cardinality} (\textbf{R2}). 
%
Finally, it presents a calculable error bound that the actual number of elements
likely falls within given a confidence probability \cite{papapetrou2010cardinality}.

@@

\autoref{f:design} illustrates the workflow of logic state coverage
measurement. Similar to edge coverage, it first records and buckets the branch
hit counts of every execution, yielding a \emph{logic state} as a result 
(\autoref{s:design:rec}). Then, it adds the recorded logic state to a 
\emph{bloom filter} that represents a set of observed logic states, by setting
the hash indices (\autoref{s:design:up}).  Finally, the measurement periodically
reads off the bloom filter to count the number of observed logic states
(\autoref{s:design:count}).
